Avoiding deforestation: The implications for global markets

### 2024-02-28 paper.Rmd
my changes?

---
title: "Avoiding deforestation: The implications for global markets"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Eric C. Davis, Maros Ivanic, and Brent Sohngen"
output: 
  bookdown::pdf_document2:
    keep_tex: TRUE
toc: FALSE
bibliography: references.bib
header-includes:
    - \usepackage{setspace}\doublespacing
abstract: "The projected growth in population and incomes as implied by various shared socioeconomic pathways is expected to create pressure to convert forestland into farmland. At the same time, the increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink. These goals are potentially incompatible as reversing the deforestation trend by afforesting cropland might result in negative market impacts such as higher food prices, especially in low agricultural productivity/high population growth scenarios. In order to better grasp the interrelationship and the magnitude of harm that might be caused by prioritizing the goal of mitigating deforestation, this analysis connects two respected global economic models: the Global Trade Analysis Project (GTAP) model and the Global Timber Model (GTM). Results from the model suggest that it is possible through fiscal policy to preserve 144.2 million hectares of forestland that otherwise would be converted to agricultural land by 2033. As to the economic price for doing so, the avoided deforestation would in most regions of the world result in less agricultural output and higher market prices. This is estimated to impact the well-being of global consumers by $92.1 billion, which translates to a global average cost of $10.6 per person in 2033."
---

@@

```{r echo = FALSE}
regLabels = c(
  "US" = 'US',
  "CHINA" = 'China',
  "BRAZIL" = 'Brazil',
  "CANADA" = 'Canada',
  "RUSSIA" = 'Russia',
  "EU ANNEX I" = 'EU I',
  "EU NON ANNEX I" = 'EU II',
  "SOUTH ASIA" = 'South Asia',
  "CENT AMER." = 'Central America',
  "RSAM" = 'Rest of Latin America',
  "SSAF" = 'Sub-Saharan Africa',
  "SE ASIA" = 'Southeast Asia',
  "OCEANIA" = 'Oceania',
  "JAPAN" = 'Japan',
  "AFME" = 'North Africa/Middle East',
  "E ASIA" = 'East Asia'
)

@@

@@

kableExtra::column_spec(
  kableExtra::column_spec(
    kableExtra::kable_styling(
      kableExtra::add_footnote(
        knitr::kable(
          t(out),
          escape = FALSE,
          format = 'latex',
          digits = 1,
          label = 'baseline-forest-change',
          caption = 'Forestland by region in 2023 and estimated change by 2033'
        ),
        label = c(
          'a' = 'Southeast Asia includes Brunei Darussalam, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore, Thailand, Timor-Leste, and Vietnam.',
          'b' = 'Oceania includes Australia, Cook Islands, Fiji, French Polynesia, Kiribati, Marshall Islands, Micronesia, Nauru, New Caledonia, New Zealand, Palau, Papua New Guinea, Samoa (Western Samoa), Solomon Islands, Tonga, Tuvalu, and Vanuatu.',
          'c' = 'Rest of South America includes Argentina, Bolivia, Czech Republic, Chile, Colombia, Ecuador, the Falkland Islands, French Guiana, Guyana, Paraguay, Peru, South Georgia, the South Sandwich Islands, Suriname, Uruguay, and Venezuela.',
          d = 'Western & Central Europe includes  Austria, Belgium, Bulgaria, Croatia, Denmark, Estonia, Finland, France, Hungary, Iceland, Italy, Ireland, Germany, Greece, Latvia, Liechtenstein, Lithuania, Luxembourg, Malta, Monaco, Montenegro, Netherlands, North Macedonia, Norway, Poland, Portugal, Romania, Serbia, Slovakia, Slovenia, Spain, Sweden, Switzerland, Turkey, and the United Kingdom.',
          e = 'Other Europe includes Albania, Armenia, Azerbaijan, Belarus, Bosnia and Herzegovina, Georgia,  Kosovo Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, and Ukraine.'
        ),
        threeparttable  = TRUE
      )
    ),
    1,
    border_left = TRUE
  ),
  5,
  border_right = TRUE
)
```

@@

kableExtra::column_spec(
  kableExtra::column_spec(
    kableExtra::kable_styling(
        knitr::kable(
          t(out),
          escape = FALSE,
          format = 'latex',
          digits = 1,
          caption = 'Percentage change in the amount of forested land in 2033 with deforestation mitigation policy in place (Source: Authors\' calculations)'
        )
    ),
    1,
    border_left = TRUE
  ),
  5,
  border_right = TRUE
)

@@

circlize::chordDiagram(changesInAgTrade, directional = TRUE, col = ifelse(changesInAgTrade>0,'green','red'),direction.type = "arrows", link.arr.type = 'big.arrow')

@@

out = rbind(out, `$\\Delta$ ($)` = worldGTAP$EV-baselineGTAP$EV)
)
rownames(out) = regLabels[gGTM[rownames(out)]]

kableExtra::column_spec(
  kableExtra::column_spec(kableExtra::kable_styling(
    knitr::kable(out, caption = 'Change in consumer well-being in 2033, by region, in dollars, with deforestation mitigation policy in place', digits = 1, escape = FALSE, label = 'EV')
  ), 1, border_left = TRUE),
  5,
  border_right = TRUE
)

@@

```{r echo= FALSE}
baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')
worldGTAP = HARr::read_SL4('asset/GTAP/run_scenario1.sl4')

out = ((1+worldGTAP$pm/100)/(1+baselineGTAP$pm/100)-1)

colnames(out) = regLabels[colnames(out)]
out = cbind(out, Total = rowSums(out))

kableExtra::column_spec(
  kableExtra::column_spec(kableExtra::kable_styling(
    knitr::kable(out, caption = 'Percent change in market prices for agricultural goods in 2033, in dollars, with deforestation mitigation policy in place', digits = 1, escape = FALSE, label = 'pm')
  ), 1, border_left = TRUE),
  5,
  border_right = TRUE
)

@@

>>>>

---
title: "Avoiding deforestation: The implications for global markets"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Eric C. Davis, Maros Ivanic, and Brent Sohngen"
output: 
  bookdown::pdf_document2:
    keep_tex: TRUE
toc: FALSE
bibliography: references.bib
header-includes:
    - \usepackage{setspace}\doublespacing
abstract: "The projected growth in population and incomes is expected to create pressure to convert forestland into farmland. At the same time, the increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink. Even though these goals are  incompatible as reversing the deforestation trend by afforesting cropland would result in negative market impacts such as higher food prices, using the GTAP and GTM models, we find that these impacts would be relatively small if the goal of  preserving 144.2 million hectares of forestland that otherwise would be converted to agricultural land by 2033 is achieved through a tax on using land in agricultural production. As to the economic price for doing so, the avoided deforestation would in most regions of the world result in less agricultural output and higher market prices. This is estimated to impact the well-being of global consumers by $92.1 billion, which translates to a global average cost of $10.6 per person in 2033."
---

@@

```{r echo = FALSE}
regLabels = c(
  "US" = 'United States',
  "CHINA" = 'China',
  "BRAZIL" = 'Brazil',
  "CANADA" = 'Canada',
  "RUSSIA" = 'Russia',
  "EU ANNEX I" = 'West/Central Europe',
  "EU NON ANNEX I" = 'EU II',
  "SOUTH ASIA" = 'South Asia',
  "CENT AMER." = 'Central America',
  "RSAM" = 'Other Latin America',
  "SSAF" = 'Sub-Saharan Africa',
  "SE ASIA" = 'SE Asia',
  "OCEANIA" = 'Oceania',
  "JAPAN" = 'Japan',
  "AFME" = 'North Africa/Middle East',
  "E ASIA" = 'East Asia'
)

@@

gGTM['ROW']='ROW'
regLabels['ROW']='ROW'

@@

kableExtra::column_spec(
  kableExtra::column_spec(
    kableExtra::kable_styling(
      kableExtra::add_footnote(
        knitr::kable(
          t(out),
          escape = FALSE,
          #format = 'latex',
          digits = 1,
          label = 'baseline-forest-change',
          caption = 'Forestland by region in 2023 and estimated change by 2033'
        ),
        label = c(
          'a' = 'Southeast Asia includes Brunei Darussalam, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore, Thailand, Timor-Leste, and Vietnam.',
          'b' = 'Oceania includes Australia, Cook Islands, Fiji, French Polynesia, Kiribati, Marshall Islands, Micronesia, Nauru, New Caledonia, New Zealand, Palau, Papua New Guinea, Samoa (Western Samoa), Solomon Islands, Tonga, Tuvalu, and Vanuatu.',
          'c' = 'Rest of South America includes Argentina, Bolivia, Czech Republic, Chile, Colombia, Ecuador, the Falkland Islands, French Guiana, Guyana, Paraguay, Peru, South Georgia, the South Sandwich Islands, Suriname, Uruguay, and Venezuela.',
          d = 'Western & Central Europe includes  Austria, Belgium, Bulgaria, Croatia, Denmark, Estonia, Finland, France, Hungary, Iceland, Italy, Ireland, Germany, Greece, Latvia, Liechtenstein, Lithuania, Luxembourg, Malta, Monaco, Montenegro, Netherlands, North Macedonia, Norway, Poland, Portugal, Romania, Serbia, Slovakia, Slovenia, Spain, Sweden, Switzerland, Turkey, and the United Kingdom.',
          e = 'Other Europe includes Albania, Armenia, Azerbaijan, Belarus, Bosnia and Herzegovina, Georgia,  Kosovo Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, and Ukraine.'
        ),
        threeparttable  = TRUE
      )
    ),
    1,
    border_left = TRUE
  ),
  5,
  border_right = TRUE
)
```

@@

kableExtra::column_spec(
  kableExtra::column_spec(
    kableExtra::kable_styling(
        knitr::kable(
          t(out),
          escape = FALSE,
          #format = 'latex',
          digits = 1,
          caption = 'Percentage change in the amount of forested land in 2033 with deforestation mitigation policy in place (Source: Authors\' calculations)'
        )
    ),
    1,
    border_left = TRUE
  ),
  5,
  border_right = TRUE
)

@@

colnames(changesInAgTrade) = rownames(changesInAgTrade) = regLabels[gGTM[c("ROW", "china", "japan", "ROW", "seasia", "ROW", "ROW", "usa", "ROW", "brazil", "samerica", "eui", "ROW", 
"ROW", "ssa", "ROW", "ROW")]]

changesInAgTrade = matsbyname::aggregate_byname(changesInAgTrade)

diag(changesInAgTrade)=0

#round(changesInAgTrade,0)

circlize::chordDiagram(changesInAgTrade, directional = TRUE, col = ifelse(changesInAgTrade>0,'green','red'),direction.type = c("arrows"), link.arr.type = 'big.arrow')

@@

out = cbind(
  `Baseline (B\\$)` = baselineGTAP$ev[, 'TOTAL'] / 1000,
  `Scenario (B\\$)` = worldGTAP$ev[, 'TOTAL'] / 1000,
  `$\\Delta$ (B\\$)` = (worldGTAP$ev[, 'TOTAL'] - baselineGTAP$ev[, 'TOTAL']) /
    1000
)

#rownames(out)[length(rownames(out))] = 'ROW'

out = rbind(out, 'Total'=colSums(out))

kableExtra::column_spec(
  kableExtra::column_spec(kableExtra::kable_styling(
    knitr::kable(out, caption = 'Change in consumer well-being in 2033, by region, in billions dollars, with deforestation mitigation policy in place', digits = 1, escape = FALSE, label = 'EV',format.args = list(big.mark = ","))
  ), 1, border_left = TRUE),
  4,
  border_right = TRUE
)

@@

```{r echo= FALSE}

out = data.frame('Ag prices'=(1+worldGTAP$pm['ag',,'TOTAL']/100)/(1+baselineGTAP$pm['ag',,'TOTAL']/100)-1, check.names = FALSE) *100

rownames(out) = regLabels[gGTM[rownames(out)]]
#out = cbind(out, Total = rowSums(out))
#out[4,'Total'] = (out[2 , 'Total']/out[1 , 'Total']-1)*100

kableExtra::column_spec(
  kableExtra::column_spec(kableExtra::kable_styling(
    knitr::kable(out, caption = 'Percent change in market prices for agricultural goods in 2033, with deforestation mitigation policy in place, relative to baseline', digits = 1, escape = FALSE, label = 'pm')
  ), 1, border_left = TRUE),
  2,
  border_right = TRUE
)

@@

# Conclusions
